### 2012-04-22 docs/blob.tex
migrating in an old, old document from the SVN repo

\documentclass[12pt]{article}
\usepackage{amssymb,amsmath,mathrsfs,color}

% functions for text
\newcommand{\foreign}[1]{\emph{#1}}
\newcommand{\project}[1]{\textsl{#1}}

% definitions for text (depend on functions)
\newcommand{\equationname}{equation}
\newcounter{hogg}
\setcounter{hogg}{0}
\renewcommand{\thehogg}{\alph{hogg}}
\newcommand{\hoggitem}{\textsl{\refstepcounter{hogg}(\thehogg)}}

% functions for math
\renewcommand{\vector}[1]{\boldsymbol{#1}}
\newcommand{\unitvector}[1]{\vector{\hat{#1}}}
\newcommand{\tensor}[1]{\boldsymbol{#1}}
\newcommand{\transpose}[1]{{#1}^{\mathsf{\!T}}}
\renewcommand{\outer}[2]{{#1}\cdot\transpose{#2}}

% definitions for math (depend on functions)
\newcommand{\allimages}{\left\{\image_n\right\}}
\newcommand{\allpars}{\vector{\theta}}
\newcommand{\band}{B}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\deflection}{\vector{\alpha}}
\newcommand{\dposition}{\vector{x}}
\newcommand{\hyperpars}{\vector{\alpha}}
\newcommand{\image}{\vector{I}}
\newcommand{\magnification}{\tensor{M}}
\newcommand{\normal}{\mathscr{N}}
\newcommand{\pars}{\vector{\omega}}
\newcommand{\phihat}{\unitvector{\phi}}
\newcommand{\potential}{\psi}
\newcommand{\residual}{\vector{e}}
\newcommand{\rhat}{\unitvector{r}}
\newcommand{\synthetic}{\vector{\tilde{I}}}
\newcommand{\variance}{\vector{\sigma}^2}

\begin{document}\sloppy\sloppypar
\section*{Is that blob a gravitational lens?}
\noindent
authors TBD

\begin{abstract}
In large ground-based imaging surveys, strong gravitational lenses are
rare and also barely resolved.  Their confident identification without
space-based imaging requires sensitive probabilistic modeling.  Here
we present the fundamentals of the probabilistic modeling approach to
finding barely resolved lenses.  Fundamentally, lenses can be found
because there is limited freedom in the possible appearance of a
source multiply imaged by a reasonable foreground mass distribution.
\end{abstract}

Under the strong assumptions that \textsl{(a)}~multiply imaged quasars
and galactic nuclei are valuable and \textsl{(b)}~extremely large
ground-based surveys will be conducted in the future, it is important
to develop methods that can identify samples of candidate
gravitational lenses with high purity and completeness.  Now that
gravitational lensing is an abundantly observed phenomenon, the
community becomes convinced that any individual source of complex
morphology is produced by gravitational lensing by being shown that
the gravitational-lens model provides a simple explanation of the
non-trivial morphology.  Projects that have convincingly demonstrated
that many sources are lenses have done so by building good (highly
likely) and reasonable (highly prior probable) models with
time-consuming gravitational lensing modeling codes (cite, for
example, slacs follow-up).

If the lens modeling could be fully automated and made fast, then in
principle samples of candidate gravitational lenses would not have to
be modeled \emph{after} their selection for study; the selection
itself could involve making good and reasonable models, and competing
them against non-lensing models.  Here ``fully automated'' means being
able to run on hundreds of millions of sources without a large number
of failures, and ``fast'' means on the order of CPU-seconds per source
(even assuming the availablility of many CPU-years of computation per
survey).

A key idea in what follows is model complexity or model freedom: When
a model is less free, it is less able to accurately represent
arbitrary sources, but it has more predictive power, because
morphologies outside its ``coverage'' are ruled out \foreign{a
  priori}.  This concept of model freedom is \emph{not} directly
related to the number of free parameters (except in extremely rare
cases).  Indeed, in some of our model comparisons below, the
gravitational lensing model will have \emph{more} free parameters but
\emph{less} model freedom than its competitor.

In what follows, in all cases, the image model and any
point-spread-function model is a model of the pixel-convolved scene,
so that interpolation of the image onto a finite pixel grid involves
only sampling and not also convolution.

\section{Model comparison generalities}

In what follows, we will think of our ``data'' as being a set of $n$
astrometrically aligned overlapping images $\image_n$, each of which
is in a known band $\band_n$ and photometrically calibrated.  We will
think of our ``model'' as being a parameterized (possibly with a large
number of parameters) prescription for generating a set of synthetic
images $\synthetic_n$, one for each image in the data set, a (possibly
parameterized) prescription for the noise-generating process in the
images that permits calculation of the probability
$p(\image_n|\synthetic_n)$ of the data given the synthetic image,
and---importantly---a description of the prior probability
distribution function (prior PDF) over all the parameters of the
synthetic images and the noise model.  That is, the model can be
described by
\begin{eqnarray}\displaystyle
\image_n &=& \synthetic_n(\allpars) + \residual_n
\\
p(\image_n|\allpars) &=& p(\residual_n|\allpars)
\\
p(\allimages) &=& \int \left[\prod_n p(\image_n|\allpars)\right]\,p(\allpars)\,\dd\allpars
\label{eq:marginalizedlikelihood}\quad,
\end{eqnarray}
where the synthetic images $\synthetic_n$ are explicitly a function of
the model parameters $\allpars$, the noise contribution to image $I_n$
is the difference between the data image and the synthetic image, the
single-image likelihood $p(\image_n|\allpars)$ is just the PDF (or,
strictly, ``frequency distribution function'') for the noise evaluated
at the residual image (difference between image and synthetic image)
$\residual_n$, the marginalized likelihood $p(\allimages)$ is the
likelihood for all images, integrated over all parameter values with
the prior (which is the measure in parameter space), and there is an
implicit independence assumption with the product of the single-image
likelihoods inside the integral.  In principle, the parameters
$\allpars$ could split into two categories: There could be primary
parameters $\pars$ and hyperparameters $\hyperpars$, with the
hyperparameters controlling aspects of the prior PDF on the
parameters.  This does not change any of the discussion here or below;
it just complexifies the enormous integral performed in
\equationname~(\ref{eq:marginalizedlikelihood}) to
\begin{eqnarray}\displaystyle
p(\allimages|\hyperpars) &=& \int \left[\prod_n p(\image_n|\pars)\right]\,p(\pars|\hyperpars)\,p(\hyperpars)\,\dd\pars
\quad .
\end{eqnarray}
Things will get more complicated when we think of the hyperpars as
applying to many lenses, and not just one at a time, but we will
postpone that for subsequent work.

Qualitatively different models can be compared by many methods.
\hoggitem~There are information criteria such as AIC (cite) and BIC
(cite), which compare the goodness of fit (likelihood) at
maximum-likelihood (that is, unmarginalized best-fit likelihood) but
penalize model complexity with a simple function of the number of
parameters.  These criteria are only strictly appropriate for linear
models (and our models are not linear), and they make strong
assumptions about parameter priors.  \hoggitem~There is
cross-validation, in which small fractions of the data are left out,
the model is fit to the remainders and used to predict the left-out
fractions.  These methods do not have a clear probabilistic basis,
although they are advocated in other places by one of us (cite hogg).
\hoggitem~There is marginalized likelihood, which is model-selection
via the scalar developed in
\equationname~(\ref{eq:marginalizedlikelihood}) above.  This
properly accounts for differences in model complexity and permits
comparisons of models without any \foreign{a priori} idea of which is
more probable.  It has the great disadvantage in practice that it
depends on having appropriately informative prior PDFs on model
parameters, a condition that is rarely met; many marginalized
likelihood calculations in the literature are extremely misleading.
\hoggitem~There is relative marginalized posterior probability, which
is the same as marginalized likelihood but multiplied by constants
(proportional to model prior probabilities) that turn the results into
marginalized posterior model probabilities.  As with the parameter
prior PDFs, when it comes to these model priors, it is ``garbage in,
garbage out''.  \hoggitem~There is expected utility maximization:
Probabilistic inference tells you what probabilities to assign, but
does not tell you what to \emph{decide}; decisions (for example the
decisions to point follow-up telescopes or spectrographs at particular
candidates) require all the likelihood and prior information mentioned
above \emph{plus} some assessment of the utilities of different
options, usually in micro-economic terms.

All of the probabilistically justified model-selection criteria in
this list (that are appropriate for non-linear models) require full
specification of the likelihood funciton and appropriately informative
priors on all model parameters.  For this reason, in what follows, we
focus on generating good marginalized likelihoods for the models we
compare, and only subsequently and briefly discuss the specific
choices for model selection in terms of marginalized likelihood,
marginalized posterior probability, and expected utility.

\section{Models}

Lens candidate identification will proceed by model comparison.
However, there are a number of components common to all models.  The
first component common to all models is the noise model.  We will
assume that each pixel in the residual image $\residual_n$---the
difference image between the data image $\image_n$ and the synthetic
image $\synthetic_n$---is drawn from a Gaussian distribution with zero
mean and variance taken from an image $\variance_n$ built by:
\begin{eqnarray}\displaystyle
\variance_n &=& a_n + b_n\,\synthetic_n
\quad ,
\end{eqnarray}
where $\variance_n$ is an image of noise variance values, one per
image pixel, and $a_n$ and $b_n$ are scalar constants, with $a_n$
relating to sky, dark, and read noise, and $b_n$ relating to
source-photon shot noise.  The constants $a_n$ and $b_n$ (two
parameters per image) will be part of the model parameters $\allpars$.
The Gaussian assumption is a large-number limit.  The only exceptions
to this independent-pixel noise model will be some masked pixels,
which will not be used in the inference at all; the mask images will
be provided by the data provider and constitute part of the assumed
prior knowledge.

The second component common to all models pixel-convolved PSF model.
We assume that a point-source can be represented in each image $n$ by
a sum of $K$ concentric two-dimensional Gaussians.  Each of these
Gaussians has mean zero but a two-dimensional variance tensor with
arbitrary one-dimensional variances and co-variance.

The third component common to all models is the synthetic galaxy
model.  Each source will be modeled as a galaxy plus one or more point
sources.  The galaxy will be parameterized by a radial model
(exponential or deVaucouleurs) with a flux, an effective (half-light)
radius, and an ellipticity tensor (axis ratio and orientation).  The
only unusual aspects to our galaxy model are the following: We choose
to truncate linearly the galaxy model intensity from the model to zero
linearly over the radius range from 7 to 8 effective radii; this is
the same choice as that made by the \project{Sloan Digital Sky Survey}
imaging pipeline (hogg cite some non-reference).  We don't use the
truncated functions (truncated exponential and truncated
deVaucouleurs) directly, but rather we use mixture-of-Gaussian
approximations to those functions.  These approximations are shown in
\figurename~HOGG.

\paragraph{$M$-image lens with no millilensing.}
The singular isothermal sphere with external shear (quadrupole
provided by a density perturbation at large radius) projects to a
two-dimensional lensing potential scalar $\potential$, deflection
angle vector $\deflection$, and magnification tensor $\magnification$
\begin{eqnarray}\displaystyle
\potential(\dposition) &=& b\,r + \frac{1}{2}\,\gamma\,r^2\,\cos(2\,(\phi-\phi_\gamma))
\\
\deflection(\dposition) &=& -b\,\rhat
 - \gamma\,r\,\cos(2\,(\phi-\phi_\gamma))\,\rhat
 + \gamma\,r\,\sin(2\,(\phi-\phi_\gamma))\,\phihat
\\
\magnification(\dposition) &=& \outer{\rhat}{\rhat} + \outer{\phihat}{\phihat} - HOGG
\\
r &\equiv& \sqrt{\dposition\cdot\dposition}
\\
\rhat &\equiv& \frac{1}{r}\,\dposition
\\
\rhat\cdot\rhat &\equiv& 1
\\
\phihat\cdot\rhat &\equiv& 0
\\
\phihat\cdot\phihat &\equiv& 1
\quad ,
\end{eqnarray}
where $\dposition$ is a two-dimensional sky vector relative to the
lens center, $b$ is the Einstein radius (units of angle), $r$ is a
radial coordinate relative to the lens center (angle units), $\gamma$
is the external shear amplitude, $\phi_\gamma$ is an orientation angle
associated with the external shear, and $\rhat$ and $\phihat$ are
unit vectors in the two-dimensional space, and implicitly the
two-dimensional vectors are unit vectors.

\paragraph{$M$-image lens with millilensing.}

\paragraph{galaxy plus central AGN}

\paragraph{galaxy plus $M$ point sources}

\section{Tests and examples}

\section{Discussion}

\end{document}